---
title: "sleep project"
author: "Allison Buck"
date: "2023-11-02"
output:
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

read in libraries and view original data set 
```{r}
library(readxl)

sleep <- read_xlsx("sleep2.xlsx")
```

view the structure and summary of the original data 
```{r}
str(sleep)
summary(sleep)
```

remove rows with empty cells and view data again 
```{r}
sleep[sleep == ''] <- NA
sleep2 <- na.omit(sleep)
```

remove unnecessary variables
```{r}
sleep2 <- subset(sleep2, select = -c(mturk, whatdisorder))
```

histograms to assess whether transformations are needed for quantitative variables 
```{r}
par(mfrow=c(2,2))
hist(sleep2$lwsleep)
hist(sleep2$lnsleep)
hist(sleep2$age)
hist(sleep2$optimalsleep)

hist(log(sleep2$age))
```

add lnage as a variable to the data set
```{r}
sleep2$lnage <- log(sleep2$age)
```

box plots for categorical variables 
```{r}
depression <- table(sleep2$depression)
depression 

anxiety <- table(sleep2$anxiety)
anxiety

epworth <- table(sleep2$epworth)
epworth

rmeq <- table(sleep2$rmeq)
rmeq

female <- table(sleep2$female)
female

minority <- table(sleep2$minority)
minority

student <- table(sleep2$student)
student

plot <- boxplot(lwsleep ~ female, data = sleep2)

boxplot(lwsleep ~ minority, data = sleep2)

boxplot(lwsleep ~ student, data = sleep2)
```

scatter plot matrix and correlations
```{r}
pairs(~ lnsleep + age + optimalsleep + lnage + lwsleep, data = sleep2, upper.panel = NULL)

cor(sleep2[, c("lnsleep", "depression", "anxiety", "epworth", "rmeq", "female", "minority", "student", "optimalsleep", "lnage", "lwsleep")], use = "complete.obs")
```

view structure and summary of cleaned data set
```{r}
str(sleep2)
summary(sleep2)
```

start of project! 

model 1: full model with not transformations
```{r}
model_1 <- lm(lwsleep ~ lnsleep + depression + anxiety + epworth + rmeq + female + minority + student + optimalsleep + sdisorder + lnage, data = sleep2)
summary(model_1)
```

model 2: all significant variables from full model 
```{r}
model_2 <- lm(lwsleep ~ lnsleep + anxiety + rmeq + epworth + optimalsleep, data = sleep2)
summary(model_2)
```

model 3: forward stepwise 
```{r}
# install packages
if(!require('MASS')) {  
	install.packages('MASS')
}
library(MASS)

# null model
null_model <- lm(lwsleep ~ 1, data = sleep2)  

# Use AIC as the selection criteria
model_3 <- stepAIC(null_model, scope = list(lower = null_model, upper = model_1), k = 2, direction = "forward")
summary(model_3)
```

model 4: backward stepwise
```{r}
model_4 <- stepAIC(model_1, scope = list(lower = null_model, upper = model_1), k = 2, direction = "backward")  
summary(model_4)
```

model 5: interaction terms -- added depression and anxiety interaction term into the full model
```{r}
# correlation coefficients between variables 
cor(sleep2[, c("lnsleep", "depression", "anxiety", "epworth", "rmeq", "female", "minority", "student", "optimalsleep", "lnage", "lwsleep")], use = "complete.obs")

fit1 <- lm(lwsleep ~ anxiety + depression, data = sleep2)
summary(fit1)

fit2 <- lm(lwsleep ~ anxiety + depression + anxiety*depression, data = sleep2)
summary(fit2)

anova(fit1, fit2)

model_5 <- lm(lwsleep ~ lnsleep + depression + anxiety + epworth + rmeq + female + minority + student + optimalsleep + sdisorder + lnage + anxiety*depression, data = sleep2)

summary(model_5)
```

model 6: add the interaction term (anxiety*depression) and depression added to model 2 (our model of significant variables based on the summary of the full)
```{r}
model_6 <- lm(formula = lwsleep ~ lnsleep + optimalsleep + anxiety + rmeq + 
    epworth + depression + anxiety*depression, data = sleep2)

summary(model_6)
```

collinearity and VIF 
```{r}
# lnsleep
VIF_lnsleep <- summary(lm(formula = lnsleep ~ optimalsleep + anxiety + rmeq + 
    epworth + depression + anxiety*depression, data = sleep2))
1/(1-VIF_lnsleep$r.squared)

# optimalsleep
VIF_optimalsleep <- summary(lm(formula = optimalsleep ~ lnsleep + anxiety + rmeq + 
    epworth + depression + anxiety*depression, data = sleep2))
1/(1-VIF_optimalsleep$r.squared)

# anxiety
VIF_anxiety <- summary(lm(formula = anxiety ~ lnsleep + optimalsleep + rmeq + 
    epworth + depression + anxiety*depression, data = sleep2))
1/(1-VIF_anxiety$r.squared)

# rmeq
VIF_rmeq <- summary(lm(formula = rmeq ~ lnsleep + optimalsleep + anxiety + 
    epworth + depression + anxiety*depression, data = sleep2))
1/(1-VIF_rmeq$r.squared)

# epworth
VIF_epworth <- summary(lm(formula = epworth ~ lnsleep + optimalsleep + anxiety + 
    rmeq + depression + anxiety*depression, data = sleep2))
1/(1-VIF_epworth$r.squared)

# depression
VIF_depression <- summary(lm(formula = depression ~ lnsleep + optimalsleep + anxiety + rmeq + 
    epworth + anxiety*depression, data = sleep2))
1/(1-VIF_depression$r.squared)

# anxiety*depression
VIF_anxiety.depression <- summary(lm(formula = anxiety*depression ~ lnsleep + 
    optimalsleep + anxiety + rmeq + epworth + depression, data = sleep2))
1/(1-VIF_anxiety.depression$r.squared)
```

anova table for model 1 and model 2 
```{r}
anova(model_1, model_2)
```

anova table for model 5 and model 6 
```{r}
anova(model_5, model_6)
```

anova table for model 2 and model 6 
```{r}
anova(model_2, model_6)
```

t-test for anxiety on model 1
```{r}
n <- 2218
k <- 11
alpha <- 0.01
qt(1-(alpha/2), n-k-1)
```


